---
title: "A Guide to the AEC's Federal Election Raw Datasets"
---
```{=html}
<style>
  .content p {
    margin-top: 30px !important;
  }

  details {
    margin-bottom: 40px;
  }
</style>
```
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE,
  warning = FALSE,
  message = FALSE
)
```
<br>
This article explores the [`get_election_data`](https://sarahcgallLtd.github.io/scgElectionsAU/reference/get_election_data.html) function and the data that it is able to retrieve from the Australian Electoral Commission's (AEC) website. This function is primarily used to retrieve the raw data with limited processing. To remove any processing, ensure the function's `process` param is set to `FALSE`. For a more processed and 'analysis-ready' function, check out [`get_results`](https://sarahcgallLtd.github.io/scgElectionsAU/reference/get_results.html).
<br>

```{r setup, include = FALSE}
library(scgElectionsAU)
library(knitr)
```
In the [`get_election_data`](https://sarahcgallLtd.github.io/scgElectionsAU/reference/get_election_data.html) function, the `type` argument default is "Federal Election". This type extracts data from the AEC's 'Tally Room' for elections between 2004 and 2022. Both House and Senate data are extractable. Additionally, when `type` is "By-Election" or "Referendum", this will extract data for by-elections between 2008 and 2024 and the 2023 referendum. There is an option for any or all types to be selected (e.g., to retrieve data for both federal elections and by-elections).

## Candidates & Nominations
### National list of candidates
The `National list of candidates` dataset contains a comprehensive list of all candidates who participated in Australian federal elections. It covers both `Senate` and `House` elections from `2004` to the present and `type=By-Election`. The `2004` dataset is processed in the following way:

- The `SittingMemberFl` column is renamed to `Elected` using the [`process_elected()`](https://docs.sarahcgall.co.uk/scgElectionsAU/reference/process_elected.html) function.
- Based on this column, the data are further standardised to match later years with "Y" and "N" values inputted. This replaces the original corresponding "#" and NA values.

This dataset enables you to:

- View candidate details (`CandidateID`, `Surname`, `GivenNm`) and their party affiliations (`PartyAb`, `PartyNm`).
- Map candidates to electoral divisions (`DivisionID`, `DivisionNm`) and states (`StateAb`).
- Identify elected candidates (`Elected`) and incumbents (`HistoricElected`).

NB. There is no consistent pattern in when candidates retain their candidate ID (`CandidateID`) or when they do not. Some candidates appear to retain theirs across elections, while others do not. It is therefore not a column to correctly identify candidates across elections. Instead it should only be used across datasets of a single election.

#### *Example Usage:*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "National list of candidates",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  type = c("Federal Election", "By-Election"), # Any or All for House only. Default option for Senate (Federal Election).
  category = "House", # OR "Senate"
)
```

#### *Sample Data:*

```{r, echo = FALSE}
# Sample data (hidden execution)
national_list_of_candidates <- suppressMessages(
  get_election_data(file_name = "National list of candidates",
  date_range = list(from = "2024-04-13", to = "2024-06-13"),
  type = c("By-Election")
))

# Select the first row for display
first_row <- national_list_of_candidates[1,]

# Display as an HTML table
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```
<br>

### Members/Senators elected
The `Members elected` and `Senators elected` datasets provide comprehensive lists of the elected Members of Parliament
(MPs) and Senators at federal elections. These datasets cover elections from `2004` to the present and are available
for the `House` and `Senate`, respectively.

- `Members elected` (`House`): Contains details of MPs who won their respective electoral divisions.
- `Senators elected` (`Senate`): Contains details of Senators elected for each state or territory, including the `ElectedOrder` column, which specifies the order in which they were elected.

There is no additional processing for these datasets, as both are already standardised across all election years.

This dataset enables you to:

- Identify elected representatives:
- For the `House`: MPs by electoral division (`DivisionID`, `DivisionNm`).
- For the `Senate`: Senators by state or territory (`StateAb`), including their election order (`ElectedOrder`).
- View candidate details (`CandidateID`, `Surname`, `GivenNm`) and party affiliations (`PartyNm`, `PartyAb`).

#### *Example Usage (Members Elected):*

```{r, eval = FALSE}
# Example usage
df <- get_election_data(
  file_name = "Members elected",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "House"
)
```

#### *Sample Data (Members Elected):*

```{r, echo = FALSE}
# Sample data (hidden execution)
members_elected <- suppressMessages(
  get_election_data(file_name = "Members elected")
)

# Select the first row for display
first_row <- members_elected[1,]

# Display as an HTML table
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```
<br>

#### *Example Usage (Senators Elected):*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Senators elected",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "Senate"
)
```

#### *Sample Data (Senators Elected):*

```{r, echo = FALSE}
senators_elected <- suppressMessages(
  get_election_data(file_name = "Senators elected", category = "Senate")
)

# Selecting only the first row and keeping all columns
first_row <- senators_elected[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```
<br>

### Nominations
The `Nominations by state` dataset provides a summary of nominations for federal elections, detailing:

- The number of divisions (`Divisions`) and nominations (`Nominations`) per state (`StateAb`, `StateNm`) for the `House`.
- The number of seats (`Seats`), candidates (`Candidates`), groups (`Groups`), and ungrouped candidates (`UngroupedCandidates`) per state (`StateAb`, `StateNm`) for the `Senate`.

Covering elections from `2004` to the present, these datasets are available for both the `House` and `Senate` categories. No additional processing is required, as the data are already standardised across all election years.

#### *Example Usage 'by state' (House):*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Nominations by state",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "House"
)
```

#### *Sample Data 'by state' (House):*

```{r, echo = FALSE}
noms_by_state <- suppressMessages(
  get_election_data(file_name = "Nominations by state")
)

# Selecting only the first row and keeping all columns
first_row <- noms_by_state[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```
<br>

#### *Example Usage 'by state' (Senate):*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Nominations by state",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "Senate"
)
```

#### *Sample Data 'by state' (Senate):*

```{r, echo = FALSE}
noms_by_state <- suppressMessages(
  get_election_data(file_name = "Nominations by state", category = "Senate")
)

# Selecting only the first row and keeping all columns
first_row <- noms_by_state[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```
<br>

The `Nominations by division` dataset provides a summary of nominations for federal elections, detailing:

- The number of nominations (`Nominations`) per division (`DivisionID`, `DivisionNm`, `DivisionAb`) for the `House`.
- Enrolment figures (`Enrolment`) and demographic classifications (`Demographics`) for each electorate.

Covering elections from `2011` to the present, these datasets are available for the `House` only. Minor amendments are made to the original dataset, changing `TransactionId` to `TransactionID` and `DivisionId` to `DivisionID`. This ensures consistency in column names across all datasets.

#### *Example Usage 'by division' (House):*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Nominations by division",
  date_range = list(from = "2011-01-01", to = "2025-01-01"),
  category = "House"
)
```

#### *Sample Data 'by division' (House):*

```{r, echo = FALSE}
noms_by_div <- suppressMessages(
  get_election_data(file_name = "Nominations by division")
)

# Selecting only the first row and keeping all columns
first_row <- noms_by_div[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```
<br>

The `Nominations by gender` dataset provides a summary of nominations for federal elections, detailing the number of nominations (`Nominations`) per party (`PartyAb`, `PartyNm`) for the `House` and `Senate`. Each state and territory have their own columns for the two genders (`MalesNSW`, `FemalesNSW`) and an unspecified column (`UnspecifiedNSW`). There is also a grand total (`GrandTotal`) of nominations by party which is useful for turning data into percentages.

Covering elections from `2004` to the present, these datasets are available for both the `House` and `Senate` categories. No additional processing is required, as the data are already standardised across all election years.

#### *Example Usage 'by gender' (House):*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Nominations by gender",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "House" # OR "Senate"
)
```

#### *Sample Data 'by gender' (House):*

```{r, echo = FALSE}
noms_by_gender <- suppressMessages(
  get_election_data(file_name = "Nominations by gender")
)

# Selecting only the first row and keeping all columns
first_row <- noms_by_gender[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

### Party representation
The `Party representation` dataset provides a summary of elected representatives for federal elections, detailing:

- The number of MPs per party (`Party`) per state and territory (`NSW`,`VIC`,`QLD`,`WA`,`SA`,`TAS`,`ACT`,`NT`) and nationally (`National`) for the `House`. It also includes the number of MPs previously representing each party at the previous election (`LastElection`).
- The number of Senators per party (`PartyAb`, `PartyNm`) per state and territory (`NSW`,`VIC`,`QLD`,`WA`,`SA`,`TAS`,`ACT`,`NT`) and nationally (`TOTAL`) for the `Senate`.

Covering elections from `2004` to the present, these datasets are available for both the `House` and `Senate` categories. In order to standardise datasets across election years, additional processing is conducted for election years prior to `2010` in the `House`. Using the [`process_reps()`](https://docs.sarahcgall.co.uk/scgElectionsAU/reference/process_reps.html) function:

- Pre-2010 columns are renamed from `Total` to `National` and from `LastElectionTotal` to `LastElection`.
- The `PartyAb` column is removed.

No additional processing is required for the `Senate`, as the data are already standardised across all election years.

#### *Example Usage (House):*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Party representation",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "House",
  process = TRUE # OR FALSE
)
```

#### *Sample Data (House):*

```{r, echo = FALSE}
party_rep <- suppressMessages(
  get_election_data(file_name = "Party representation")
)

# Selecting only the first row and keeping all columns
first_row <- party_rep[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```
<br>

#### *Example Usage (Senate):*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Party representation",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "Senate"
)
```

#### *Sample Data (Senate):*

```{r, echo = FALSE}
party_rep <- suppressMessages(
  get_election_data(file_name = "Party representation", category = "Senate")
)

# Selecting only the first row and keeping all columns
first_row <- party_rep[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```
<br>

## Seats
### Seats that changed hands
The `Seats that changed hands` dataset provides a summary of the seats that have changed from one party to another, detailing:

- The division details (`StateAb`, `DivisionID`, `DivisionNm`).
- The previous party that held the seat (`PreviousPartyAb`, `PreviousPartyNm`).
- The details of the newly elected MP who won that seat at the election (`CadidateID`, `Surname`, `GivenNm`, `PartyAb`, `PartyNm`).

Covering elections from `2004` to the present, these datasets are available for the `House` only. No additional processing is required, as the data are already standardised across all election years.

#### *Example Usage:*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Seats that changed hands",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "House"
)
```

#### *Sample Data:*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Seats that changed hands")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

### Seats decided on FPTP or preferences
The `Seats decided on first preferences` dataset provides a summary of the seats that were decided on first preferences (i.e., the winning candidate in a division achieved equal to or more than 50% of the primary vote - similar to First-Past-The-Post), detailing:

- The division details (`StateAb`, `DivisionID`, `DivisionNm`).
- The winning party (`PartyAb`, `PartyNm`).

Covering elections from `2004` to the present, these datasets are available for the `House` only. No additional processing is required, as the data are already standardised across all election years.

#### *Example Usage 'first':*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Seats decided on first preferences",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "House"
)
```

#### *Sample Data 'first':*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Seats decided on first preferences")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```
<br>

The `Seats decided on preferences` dataset provides a summary of the seats that were decided on preferences (i.e., no candidate in a division achieved more than 50% of the primary vote), detailing:

- The division details (`StateAb`, `DivisionID`, `DivisionNm`).
- The winning party (`PartyAb`, `PartyNm`).

Covering elections from `2004` to the present, these datasets are available for the `House` only. No additional processing is required, as the data are already standardised across all election years.

#### *Example Usage 'preference':*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Seats decided on preferences",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "House"
)
```

#### *Sample Data 'preference':*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Seats decided on preferences")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```


<br>

### Non-classic divisions
The `Non-classic divisions` dataset provides a summary of the seats in which the two-party contest included a non-major party candidate, detailing:

- The division details (`StateAb`, `DivisionID`, `DivisionNm`).
- The latest time and date that the data were updated (`Updated`).
- The first candidate details (`PartyAb1`, `PartyNm1`, `CandidateID1`, `Surname1`, `GivenNm1`, `Elected1`, `HistoricElected1`).
- The second candidate details (`PartyAb2`, `PartyNm2`, `CandidateID2`, `Surname2`, `GivenNm2`, `Elected2`, `HistoricElected2`).

Covering elections from `2010` to the present, these datasets are available for the `House` only. Minor amendments are made to the original dataset, changing `CandidateId` to `CandidateID` and `DivisionId` to `DivisionID`. This ensures consistency in column names across all datasets.

#### *Example Usage:*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Non-classic divisions",
  date_range = list(from = "2010-01-01", to = "2025-01-01"),
  category = "House"
)
```

#### *Sample Data:*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Non-classic divisions")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

## First Preferences
### First preferences by party
The `First preferences by party` dataset provides a summary of the first preferences received by each party, detailing:

- The party details (`PartyAb`, `PartyNm`).
- The number of divisions that the party stood a candidate (`Contested`), the number of candidate nominations submitted by the party (`Nominations`), and the number of candidates who were elected/won their division (`Elected`).
- The number of votes that the party received by vote type (`OrdinaryVotes`, `AbsentVotes`, `ProvisionalVotes`, `DeclarationPrePollVotes`, `PostalVotes`, `TotalVotes`).
- The percentage of votes received by each party by vote type (`OrdinaryPercentage`, `AbsentPercentage`, `ProvisionalPercentage`, `PrePollPercentage`, `PostalPercentage`, `TotalPercentage`).
- The overall swing in total votes received by the party compared to the previous election (`TotalSwing`).

Covering elections from `2004` to the present, these datasets are available for the `House` only. `PrePollVotes` is renamed `DeclarationPrePollVotes` for consistency and clarification.

#### *Example Usage 'by party' (House):*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "First preferences by party",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "House"
)
```

#### *Sample Data 'by party' (House):*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "First preferences by party")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

The `First preferences by state by party` dataset further breaks down the above figures by state (`StateAb`).

#### *Example Usage of 'by state by party' (House):*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "First preferences by state by party",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "House"
)
```

#### *Sample Data of 'by state by party' (House):*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "First preferences by state by party")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

The `First preferences by group by vote type` dataset is the Senate-equivalent and provides first preference vote totals by party groupings, detailing:

- The party details (`PartyAb`, `PartyNm`).
- The number of votes that the party received by vote type (`OrdinaryVotes`, `AbsentVotes`, `ProvisionalVotes`, `DeclarationPrePollVotes`, `PostalVotes`, `TotalVotes`).
- The percentage of votes received by each party by vote type (`OrdinaryPercentage`, `AbsentPercentage`, `ProvisionalPercentage`, `DeclarationPrePollPercentage`, `PostalPercentage`, `TotalPercentage`).

Covering elections from `2004` to the present, these datasets are available for the `Senate` only. The `2004` dataset is processed in the following way:

- The `SittingMemberFl` column is renamed to `Elected` using the [`process_elected()`](https://docs.sarahcgall.co.uk/scgElectionsAU/reference/process_elected.html) function.
- Based on this column, the data are further standardised to match later years with "Y" and "N" values inputted. This replaces the original corresponding "#" and NA values.

#### *Example Usage 'by group' (Senate):*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "First preferences by group by vote type",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "Senate",
  process = TRUE # OR FALSE
)
```

#### *Sample Data 'by group' (Senate):*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "First preferences by group by vote type", category = "Senate")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

The `First preferences by state by group by vote type` dataset further breaks down the above figures by state (`StateAb`).

#### *Example Usage 'by state by group' (Senate):*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "First preferences by state by group by vote type",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "Senate",
  process = TRUE # OR FALSE
)
```

#### *Sample Data 'by state by group' (Senate):*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "First preferences by state by group by vote type", category = "Senate")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

### First preferences by candidate by vote type
The `First preferences by candidate by vote type` dataset provides a summary of the first preferences received by each individual candidate, detailing:

- The division details (`StateAb`, `DivisionID`, `DivisionNm`).
- The candidate details (`CandidateID`, `Surname`, `GivenNm`), including their position on the ballot position (`BallotPosition`), and whether they were elected (`Elected`) or an incumbent (`HistoricElected`), and the party (`PartyAb`, `PartyNm`) that they represent.
- The number of votes that the candidate received by vote type (`OrdinaryVotes`, `AbsentVotes`, `ProvisionalVotes`, `DeclarationPrePollVotes`, `PostalVotes`, `TotalVotes`).
- The overall swing in total votes received by the candidate compared to the previous election (`TotalSwing`).

Covering elections from `2004` to the present, this dataset is available for the `House` only. The `2004` dataset is processed in the following way:

- The `SittingMemberFl` column is renamed to `Elected` using the [`process_elected()`](https://docs.sarahcgall.co.uk/scgElectionsAU/reference/process_elected.html) function.
- Based on this column, the data are further standardised to match later years with "Y" and "N" values inputted. This replaces the original corresponding "#" and NA values.
- `PrePollVotes` is renamed `DeclarationPrePollVotes` for consistency and clarification.

#### *Example Usage (House):*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "First preferences by candidate by vote type",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "House",
  process = TRUE # OR FALSE
)
```

#### *Sample Data (House):*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "First preferences by candidate by vote type")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

The `First preferences by state by vote type` dataset is the equivalent for the Senate, albeit it differs in some columns. The dataset details:

- The division details (`StateAb`, `DivisionID`, `DivisionNm`).
- The candidate details (`CandidateID`), including their ballot details (`BallotPosition`,`CandidateDetails`, `Group`), and whether they were elected (`Elected`) or an incumbent (`HistoricElected`), and the party (`PartyAb`, `PartyNm`) that they represent.
- The number of votes that the candidate received by vote type (`OrdinaryVotes`, `AbsentVotes`, `ProvisionalVotes`, `DeclarationPrePollVotes`, `PostalVotes`, `TotalVotes`).

Covering elections from `2004` to the present, this dataset is available for the `Senate` only. Minor amendments are made to the original dataset, changing `PartyName` to `PartyNm` and `PrePollVotes` to `DeclarationPrePollVotes`. This ensures consistency in column names across all datasets.

#### *Example Usage (Senate Division-level):*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "First preferences by division by vote type",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "Senate"
)
```

#### *Sample Data (Senate Division-level):*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "First preferences by division by vote type", category = "Senate")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

The state-level version of this `Senate` dataset is also available by calling `First preferences by state by vote type`.

#### *Example Usage (Senate State-level):*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "First preferences by state by vote type",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "Senate",
  process = TRUE # OR FALSE
)
```

#### *Sample Data (Senate State-level):*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "First preferences by state by vote type", category = "Senate")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

### First preferences by candidate by polling place
The `First preferences by candidate by polling place` dataset provides first preference vote totals of each candidate by polling place, detailing:

- The polling place details (`StateAb`, `DivisionID`, `DivisionNm`, `PollingPlaceID`, `PollingPlaceNm`).
- The candiate details (`CandidateID`, `Surname`, `GivenNm`), including their position on the ballot position (`BallotPosition`), and whether they were elected (`Elected`) or an incumbent (`HistoricElected`), and the party (`PartyAb`, `PartyNm`) they represent.
- The number of votes that the party received (`OrdinaryVotes`).
- The overall swing in ordinary votes received by the candidate compared to the previous election (`Swing`).

Covering elections from `2004` to the present, these datasets are available for the `House` only. The `2004` dataset is processed in the following way:

- The `SittingMemberFl` column is renamed to `Elected` using the [`process_elected()`](https://docs.sarahcgall.co.uk/scgElectionsAU/reference/process_elected.html) function.
- Based on this column, the data are further standardised to match later years with "Y" and "N" values inputted. This replaces the original corresponding "#" and NA values.

*Caution should be taken if retrieving multiple election years as this dataset is large. Sizes include:

- 2022: 17 columns by 76,452 rows (1,299,684 cells)
- 2019: 17 columns by 71,805 rows (1,220,685 cells)
- 2016: 17 columns by 64,112 rows (1,089,904 cells)
- 2013: 17 columns by 82,398 rows (1,400,766 cells)
- 2010: 17 columns by 60,663 rows (1,031,271 cells)
- 2007: 17 columns by 66,269 rows (1,126,573 cells)
- 2004: 16 columns by 69,098 rows (1,105,568 cells)

#### *Example Usage:*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "First preferences by candidate by polling place",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  type = c("Federal Election", "By-Election") # Any or All
)
```

#### *Sample Data:*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "First preferences by candidate by polling place")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

## Two Party Preferred
### Two party preferred by state
The `Two party preferred by state` dataset provides the two party preferred (TPP) vote by state (`StateAb`) detailing:

- The Coalition votes and percentage (`Liberal/National Coalition Votes`, `Liberal/National Coalition Percentage`)
- The ALP votes and percentage (`Australian Labor Party Votes`, `Australian Labor Party Percentage`.
- The total votes (`TotalVotes`) and swing (`Swing`).

Covering elections from `2004` to the present, this dataset is available for the `House` only. No changes are made.

#### *Example Usage 'by state':*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Two party preferred by state",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "House"
)
```

#### *Sample Data 'by state':*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Two party preferred by state")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```
<br>

The `Two party preferred by division` dataset further breaks down the above figures by division (`DivisionID`, `DivisionNm`) and includes the winning party of the seat (`PartyAb`).

#### *Example Usage 'by division':*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Two party preferred by division",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "House"
)
```

#### *Sample Data 'by division':*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Two party preferred by division")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

### Two party preferred by division by vote type
The `Two party preferred by division by vote type` dataset provides the two party preferred (TPP) vote by vote type detailing:

- The division details (`StateAb`, `DivisionNm`, `DivisionID`).
- The winning party of the seat (`PartyAb`).
- The Coalition votes by vote type (`Liberal/National Coalition OrdinaryVotes`,`Liberal/National Coalition AbsentVotes`,`Liberal/National Coalition ProvisionalVotes`,`Liberal/National Coalition PostalVotes`,`Liberal/National Coalition DeclarationPrePollVotes`).
- The Coalition percentage by vote type (`Liberal/National Coalition OrdinaryPercentage`,`Liberal/National Coalition AbsentPercentage`,`Liberal/National Coalition ProvisionalPercentage`,`Liberal/National Coalition PostalPercentage`,`Liberal/National Coalition DeclarationPrePollPercentage`).
- The ALP votes by vote type (`Australian Labor Party OrdinaryVotes`,`Australian Labor Party AbsentVotes`,`Australian Labor Party ProvisionalVotes`,`Australian Labor Party PostalVotes`,`Australian Labor Party DeclarationPrePollVotes`).
- The ALP percentage by vote type (`Australian Labor Party OrdinaryPercentage`,`Australian Labor Party AbsentPercentage`,`Australian Labor Party ProvisionalPercentage`,`Australian Labor Party PostalPercentage`,`Australian Labor Party DeclarationPrePollPercentage`).
- The total number of votes (`TotalVotes`).

Covering elections from `2004` to the present, this dataset is available for the `House` only. No changes are made.

#### *Example Usage 'by division':*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Two party preferred by division by vote type",
  date_range = list(from = "2016-01-01", to = "2025-01-01"),
  category = "House"
)
```

#### *Sample Data 'by division':*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Two party preferred by division by vote type")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

The `Two party preferred by polling place` dataset further breaks down the above figures by polling place (`PollingPlaceID`, `PollingPlaceNm`).

#### *Example Usage 'by polling place':*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Two party preferred by polling place",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  type = c("Federal Election", "By-Election"), # Any or All
  category = "House"
)
```

#### *Sample Data 'by polling place':*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Two party preferred by polling place")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

## Two Candidate Preferred
### Two candidate preferred by candidate by vote type
The `Two candidate preferred by candidate by vote type` dataset provides details of the two candidate preferred (TCP) votes by vote type for each candidate, detailing:

- The division details (`StateAb`, `DivisionID`, `DivisionNm`).
- The candidate details (`CandidateID`, `Surname`, `GivenNm`), including their position on the ballot position (`BallotPosition`), and whether they were elected (`Elected`) or an incumbent (`HistoricElected`), and the party (`PartyAb`, `PartyNm`) that they represent.
- The number of votes that the candidate received by vote type (`OrdinaryVotes`, `AbsentVotes`, `ProvisionalVotes`, `DeclarationPrePollVotes`, `PostalVotes`, `TotalVotes`).
- The overall swing in total votes received by the candidate compared to the previous election (`Swing`).

Covering elections from `2004` to the present, this dataset is available for the `House` only. The `2004` dataset is processed in the following way:

- The `SittingMemberFl` column is renamed to `Elected` using the [`process_elected()`](https://docs.sarahcgall.co.uk/scgElectionsAU/reference/process_elected.html) function.
- Based on this column, the data are further standardised to match later years with "Y" and "N" values inputted. This replaces the original corresponding "#" and NA values.
- `PrePollVotes` is renamed `DeclarationPrePollVotes` for consistency and clarification.

#### *Example Usage:*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Two candidate preferred by candidate by vote type",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "House",
  process = TRUE # OR FALSE
)
```

#### *Sample Data:*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Two candidate preferred by candidate by vote type")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

### Two candidate preferred by candidate by polling place
The `Two candidate preferred by candidate by polling place` dataset provides details of the two candidate preferred (TCP) votes for each candidate and the polling place-level, detailing:

- The polling place details (`StateAb`, `DivisionID`, `DivisionNm`, `PollingPlaceID`, `PollingPlaceNm`).
- The candidate details (`CandidateID`, `Surname`, `GivenNm`), including their position on the ballot position (`BallotPosition`), and whether they were elected (`Elected`) or an incumbent (`HistoricElected`), and the party (`PartyAb`, `PartyNm`) that they represent.
- The number of ordinary votes that the candidate received (`OrdinaryVotes`).
- The overall swing in total votes received by the candidate compared to the previous election (`Swing`).

Covering elections from `2004` to the present, this dataset is available for the `House` only. The `2004` dataset is processed in the following way:

- The `SittingMemberFl` column is renamed to `Elected` using the [`process_elected()`](https://docs.sarahcgall.co.uk/scgElectionsAU/reference/process_elected.html) function.
- Based on this column, the data are further standardised to match later years with "Y" and "N" values inputted. This replaces the original corresponding "#" and NA values.

#### *Example Usage:*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Two candidate preferred by candidate by polling place",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  type = c("Federal Election", "By-Election"), # Any or All
  category = "House",
  process = TRUE # OR FALSE
)
```

#### *Sample Data:*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Two candidate preferred by candidate by polling place")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

## Distribution of Preferences
### Distribution of preferences by candidate
The `Distribution of preferences by candidate by division` dataset provides details of the distribution of preferences by division, detailing:

- The division details (`StateAb`, `DivisionID`, `DivisionNm`).
- The candidate details (`CandidateID`, `Surname`, `GivenNm`), including their position on the ballot position (`BallotPosition`), and whether they were elected (`Elected`) or an incumbent (`HistoricElected`), and the party (`PartyAb`, `PartyNm`) that they represent.
- The distrbution details (`CountNumber`, `CalculationType`, `CalculationValue`).

Covering elections from `2004` to the present, this dataset is available for the `House` only. The `2004` dataset is processed in the following way:

- The `SittingMemberFl` column is renamed to `Elected` using the [`process_elected()`](https://docs.sarahcgall.co.uk/scgElectionsAU/reference/process_elected.html) function.
- Based on this column, the data are further standardised to match later years with "Y" and "N" values inputted. This replaces the original corresponding "#" and NA values.

#### *Example Usage 'by division':*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Distribution of preferences by candidate by division",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "House",
  process = TRUE # OR FALSE
)
```

#### *Sample Data 'by division':*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Distribution of preferences by candidate by division")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

The `Distribution of preferences by polling place` dataset further breaks down the above figures by polling place (`PollingPlaceID`, `PollingPlaceNm`). Additional amendments are made to the original dataset, changing `CountNum` to `CountNumber`, `PPId` to `PollingPlaceID`,  `PPNm` to `PollingPlaceNm`, `CandidateId` to `CandidateID`, and `SittingMemberFl` to `Elected`. This ensures consistency in column names across all datasets.

*Caution should be taken if retrieving multiple election years as this dataset is large. Sizes include:

- 2022: 17 columns by 2,132,816 rows (36,257,872 cells)
- 2019: 17 columns by 1,719,736 rows (29,235,512 cells)
- 2016: 17 columns by 1,475,704 rows (25,086,968 cells)
- 2013: 17 columns by 2,335,760 rows (39,707,920 cells)
- 2010: 17 columns by 1,155,304 rows (19,640,168 cells)
- 2007: 17 columns by 1,597,048 rows (27,149,816 cells)
- 2004: 17 columns by 1,770,184 rows (30,093,128 cells)
- Max (All Elections + By-Elections): 17 columns by 12,835,444 rows (218,202,548 cells)

#### *Example Usage 'by polling place':*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Distribution of preferences by polling place",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  type = c("Federal Election", "By-Election"), # Any or All
  category = "House",
  process = TRUE # OR FALSE
)
```

#### *Sample Data 'by polling place':*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Distribution of preferences by polling place")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

## Flow of Preferences
### Two party preferred flow of preferences by party
The `Two party preferred flow of preferences by party` dataset provides the TPP flow of preferences, detailing:

- the party details from whom the votes are transferred (`PartyAb`,`PartyNm`).
- The votes and percentage of votes transferred to the Coalition (`Liberal/National Coalition Transfer Votes`, `Liberal/National Coalition Tansfer Percentage`)
- The votes and percentage of votes transferred to the ALP (`Australian Labor Party Transfer Votes`, `Australian Labor Party Transfer Percentage`.

Covering elections from `2004` to the present, this dataset is available for the `House` only. No changes are made.

#### *Example Usage 'by party':*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Two party preferred flow of preferences by party",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "House"
)
```

#### *Sample Data 'by party':*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Two party preferred flow of preferences by party")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

The `Two party preferred flow of preferences by state by party` dataset further breaks down the above figures by state (`StateAb`).

#### *Example Usage 'by state':*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Two party preferred flow of preferences by state by party",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "House"
)
```

#### *Sample Data 'by state':*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Two party preferred flow of preferences by state by party")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

### Two candidate preferred flow of preferences by party
The `Two candidate preferred flow of preferences by party` dataset provides the TCP flow of preferences, detailing:

- the party details from whom the votes are transferred (`FromPartyGroupID`,`FromPartyGroupAb`,`FromPartyGroupNm`,`FromPartyOrderOfAppearance`).
- the party details to whom the votes are transferred (`ToPartyGroupID`,`ToPartyGroupAb`,`ToPartyGroupNm`,`ToPartyOrderOfAppearance`).
- The number of votes and percentage of votes transferred (`TransferVotes`, `TransferPercentage`)

Covering elections from `2004` to the present, this dataset is available for the `House` only. No changes are made.

#### *Example Usage 'by party':*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Two candidate preferred flow of preferences by party",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "House"
)
```

#### *Sample Data 'by party':*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Two candidate preferred flow of preferences by party")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

The `Two candidate preferred flow of preferences by state by party` dataset further breaks down the above figures by state (`StateAb`).

*Caution should be taken if retrieving multiple election years as this dataset is large. Sizes include:

- 2022: 21 columns by 127,404 rows (2,675,484 cells)
- 2019: 21 columns by 115,350 rows (2,422,350 cells)
- 2016: 21 columns by 101,664 rows (2,134,944 cells)
- 2013: 21 columns by 136,516 rows (2,866,836 cells)
- 2010: 21 columns by 91,136 rows (1,913,856 cells)
- 2007: 21 columns by 106,838 rows (2,243,598 cells)
- 2004: 21 columns by 112,788 rows (2,368,548 cells)

#### *Example Usage 'by state':*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Two candidate preferred flow of preferences by state by party",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "House"
)
```

#### *Sample Data 'by state':*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Two candidate preferred flow of preferences by state by party")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

### Two candidate preferred flow of preferences by candidate
The `Two candidate preferred flow of preferences by candidate by division` dataset provides the TCP flow of preferences by candidate, detailing:

- the division details (`StateAb`, `DivisionID`, `DivisionNm`).
- the candidate details from whom the votes are transferred (`FromCandidateID`,`FromCandidatePartyAb`,`FromCandidatePartyNm`,`FromCandidateSurname`,`FromCandidateGivenNm`, `FromCandidateBallotPosition`).
- the candidate details to whom the votes are transferred (`ToCandidateID`,`ToCandidatePartyAb`,`ToCandidatePartyNm`,`ToCandidateSurname`,`ToGivenNm`, `ToCandidateBallotPosition`).
- The number of votes and percentage of votes transferred (`TransferVotes`, `TransferPercentage`)

Covering elections from `2004` to the present, this dataset is available for the `House` only. No changes are made.

#### *Example Usage 'by division':*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Two candidate preferred flow of preferences by candidate by division",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "House"
)
```

#### *Sample Data 'by division':*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Two candidate preferred flow of preferences by candidate by division")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

The `Two candidate preferred flow of preferences by polling place` dataset further breaks down the above figures by polling place (`PollingPlaceID`,`PollingPlaceNm`).

#### *Example Usage 'by polling place':*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Two candidate preferred flow of preferences by polling place",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  type = c("Federal Election", "By-Election"),
  category = "House"
)
```

#### *Sample Data 'by polling place':*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Two candidate preferred flow of preferences by polling place")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

## Votes
### Informal votes
The `Informal votes by state` dataset provides a summary of formal and informal votes, detailing:

- the state details (`StateAb`, `StateNm`).
- the number of formal and informal votes (`FormalVotes`,`InformalVotes`).
- the total number of votes (formal + informal) (`TotalVotes`).
- The percentage of informal votes (`InformalPercent`).
- the overall swing in informal votes received compared to the previous election (`InformalSwing`).

Covering elections from `2004` to the present, this dataset is available for both the `House`, `Senate`, or `Referendum`. No changes are made.

#### *Example Usage 'by state':*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Informal votes by state",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  # type = "Referendum", (for Referendum)
  category = "House" # OR "Senate" OR "Referendum"
)
```

#### *Sample Data 'by state':*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Informal votes by state")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

The `Informal votes by division` dataset further breaks down the above figures by division (`DivisionID`,`DivisionNm`).

#### *Example Usage 'by division':*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Informal votes by division",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  # type = "Referendum", (for Referendum)
  category = "House" # OR "Senate" OR "Referendum"
)
```

#### *Sample Data 'by division':*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Informal votes by division")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

### Turnout
The `Turnout by state` dataset provides a summary of voter turnout, detailing:

- the state details (`StateAb`, `StateNm`).
- the number of voters enrolled (`Enrolment`).
- the number and percentage of voters who voted (`Turnout`, `TurnoutPercentage`).
- the overall swing in turnout compared to the previous election (`TurnoutSwing`).

Covering elections from `2004` to the present, this dataset is available for both the `House`, `Senate`, or `Referendum`. No changes are made.

#### *Example Usage 'by state':*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Turnout by state",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  # type = "Referendum", (for Referendum)
  category = "House" # OR "Senate" OR "Referendum"
)
```

#### *Sample Data 'by state':*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Turnout by state")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

The `Turnout by division` dataset further breaks down the above figures by division (`DivisionID`,`DivisionNm`).

#### *Example Usage 'by division':*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Turnout by division",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  # type = "Referendum", (for Referendum)
  category = "House" # OR "Senate" OR "Referendum"
)
```

#### *Sample Data 'by division':*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Turnout by division")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

### Votes Type
The `Votes by state` dataset provides the number of votes by vote type, detailing:

- the state details (`StateAb`, `StateNm`).
- the number of voters enrolled (`Enrolment`).
- the number votes by vote type (`OrdinaryVotes`, `AbsentVotes`, `ProvisionalVotes`, `DeclarationPrePollVotes`, `PostalVotes`).
- the total number of votes and percentage (`TotalVotes`, `TotalPercentage`).

Covering elections from `2004` to the present, this dataset is available for the `House`, `Senate` or `Referendum`. Minor amendments are made to the original dataset, changing `PrePollVotes` to `DeclarationPrePollVotes`. This ensures consistency in column names across all datasets.

#### *Example Usage 'by state':*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Votes by state",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  # type = "Referendum", (for Referendum)
  category = "House" # OR "Senate" OR "Referendum"
)
```

#### *Sample Data 'by state':*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Votes by state")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

The `Votes by division` dataset further breaks down the above figures by division (`DivisionID`,`DivisionNm`).

#### *Example Usage 'by division':*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Votes by division",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  # type = "Referendum", (for Referendum)
  category = "House" # OR "Senate" OR "Referendum"
)
```

#### *Sample Data 'by division':*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Votes by division")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

The `Votes by polling place` dataset further breaks down the above figures by division (`PollingPlaceID`,`PollingPlaceNm`). It is only for `Referendum`.

#### *Example Usage 'by polling place':*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Votes by polling place",
  date_range = list(from = "2022-01-01", to = "2025-01-01"),
  type = "Referendum",
  category = "Referendum"
)
```

#### *Sample Data 'by polling place':*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Votes by polling place",
               type = "Referendum",
               category = "Referendum")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

## General
### Political parties
The `Political parties` dataset provides a summary of the political parties being represented at each election, detailing:

- the state details (`StateAb`).
- the party details (`PartyAb`, `PartyNm`), including the registered party name (`RegisteredPartyAb`).

Covering elections from `2004` to the present, this dataset is available for the `General` category. No changes are made.

#### *Example Usage:*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Political parties",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  category = "General"
)
```

#### *Sample Data:*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Political parties",
               category = "General")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

### Polling places
The `Polling places` dataset provides a summary of the polling places used at each election, detailing:

- the division details (`StateAb`, `DivisionID`, `DivisionNm`).
- the polling place details (`PollingPlaceID`, `PollingPlaceTypeID`, `PollingPlaceNm`).
- the address of the polling place (`PremisesAddress1`, `PremisesAddress2`, `PremisesAddress3`, `PremisesSubrub`, `PremisesStateAb`, `PremisesPostCode`).
- the coordinates of the polling place (`Latitude`, `Longitude`).

Covering elections from `2004` to the present, this dataset is available for the `General` category. The dataset is processed in the following way, using the [`process_coords()`](https://docs.sarahcgall.co.uk/scgElectionsAU/reference/process_coords.html) function:

- `NA` or `0` values in `Latitude` and `Longitude` are replaced if known, based on matching of the `PollingPlaceID` column within an internal dataset and the dataset retrieved from the AEC.

#### *Example Usage:*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Polling places",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  type = c("Federal Election", "Referendum", "By-Election"), # Any or All
  category = "General",
  process = TRUE # OR FALSE
)
```

#### *Sample Data:*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Polling places",
               category = "General")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

### Enrolment
The `Enrolment by state` dataset provides a summary of the enrolment, detailing:

- the state details (`StateAb`, `StateNm`).
- the number of enrolments based on type (`CloseOfRollsEnrolment`,`NotebookRollAdditions`,`NotebookRollDeletions`,`ReinstatementsPrePoll`,`ReinstatementsAbsent`,`ReinstatementsProvisional`).
- the overall number of enrolled voters (`Enrolment`).

Covering elections from `2004` to the present, this dataset is available for the `General` category. No changes are made.

#### *Example Usage 'by state':*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Enrolment by state",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  type = c("Federal Election", "Referendum"), # Any or All
  category = "General"
)
```

#### *Sample Data 'by state':*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Enrolment by state",
               category = "General")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

The `Enrolment by division` dataset further breaks down the above figures by division (`DivisionID`,`DivisionNm`).

#### *Example Usage 'by division':*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Enrolment by division",
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  type = c("Federal Election", "Referendum"), # Any or All
  category = "General"
)
```

#### *Sample Data 'by division':*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Enrolment by division",
               category = "General")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

### Declaration votes
The `Declaration votes issued by state` and `Declaration votes received by state` datasets provide a summary of the number of declaration votes issued and received, respectively, detailing:

- the state details (`StateAb`, `StateNm`).
- the number of declaration votes by type (`Provisional`,`Absent`,`PrePoll`,`PrePollOwnDivision`,`Postal`,`PostalOwnDivision`).

Covering elections from `2004` to the present, this dataset is available for the `General` category. No changes are made.

#### *Example Usage 'by state':*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Declaration votes issued by state", # OR Declaration votes received by state
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  type = c("Federal Election", "Referendum"), # Any or All
  category = "General"
)
```

#### *Sample Data 'by state':*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Declaration votes issued by state",
               category = "General")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

The `Declaration votes issued by division` and `Declaration votes received by division` datasets further break down the above figures by division (`DivisionID`,`DivisionNm`).

#### *Example Usage 'by division':*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Declaration votes issued by division", # OR Declaration votes received by division
  date_range = list(from = "2004-01-01", to = "2025-01-01"),
  type = c("Federal Election", "Referendum"), # Any or All
  category = "General"
)
```

#### *Sample Data 'by division':*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Declaration votes issued by division",
               category = "General")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br

## Statistics
### Postal vote applications by total
The `Postal vote applications by total` dataset provides a summary of the number of postal vote applications received and postal votes returned, detailing:

- the division details (`StateAb`, `DivisionNm`).
- the number of valid applications received (`Valid Applications Received`).
- the number of postal votes returned (`Postal Votes Returned`).

Covering the`2022` election only, this dataset is available for the `Statistics` category. No changes are made.

#### *Example Usage:*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Postal vote applications by total",
  date_range = list(from = "2022-01-01", to = "2025-01-01"),
  type = c("Federal Election", "Referendum", "By-Election"), # Any or All
  category = "Statistics"
)
```

#### *Sample Data:*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Postal vote applications by total",
               date_range = list(from = "2022-01-01", to = "2025-01-01"),
               category = "Statistics")
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

### Postal vote applications by date
The `Postal vote applications by date` dataset provides a summary of the number of postal votes returned each day, detailing:

- the division details (`StateAb`, `DivisionNm`).
- the date the postal vote was received (`DateReceived`).
- the total number of postal votes received (`TotalPVAs`).

Covering elections from `2010` to `2019`, this dataset is available for the `Statistics` category. The dataset is processed in the following way, using the [`process_pva_date()`](https://docs.sarahcgall.co.uk/scgElectionsAU/reference/process_pva_date.html) function:

- column names are standardised, including `Enrolment`, `Enrolment Division`, `PVA_Web_2_Date_Div` and `PVA_Web_2_V2_Div` are renamed to `DivisionNm` and `State_Cd` to `StateAb`.
- Unnecessary columns are dropped, including `<>`, `TOTAL to date(Inc GPV)`, and `Date out of range`.
- the date columns are pivoted from wide to long and converted into a date format for ease of comparison across years.

#### *Example Usage:*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Postal vote applications by date",
  date_range = list(from = "2010-01-01", to = "2022-01-01"),
  type = c("Federal Election", "By-Election"), # Any or All
  category = "Statistics",
  process = TRUE # OR FALSE
)
```

#### *Sample Data:*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Postal vote applications by date",
               date_range = list(from = "2019-01-01", to = "2022-01-01"),
               category = "Statistics", process = TRUE)
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

### Postal vote applications by party
The `Postal vote applications by party` dataset provides a summary of the number of postal votes returned each day, detailing:

- the division details (`StateAb`, `DivisionNm`).
- the number of General Postal Voters (`GPV`) - these are voters who automatically receive ballot papers via post and generally include voters with a disability, silent electors, prisoners, those in remote areas, and those who are unable to attend a polling place on election day due to religious objections.
- the number of AEC-received postal vote applications and by the method in which they were received (`AEC (Online)`, `AEC (Paper)`, `AEC (Total)`).
- the number of party-received postal vote applications (`ALP`, `CLP`, `DEM`, `LIB`, `LNP`, `NAT`, `OTH`).
- the total number of AEC and party received postal vote applications (`Total (AEC + Parties)`).

Covering elections from `2010` to `2019`, this dataset is available for the `Statistics` category. The dataset is processed in the following way, using the [`process_pva_party()`](https://docs.sarahcgall.co.uk/scgElectionsAU/reference/process_pva_party.html) function:

- column names are standardised, including abbreviating party names.

#### *Example Usage:*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Postal vote applications by party",
  date_range = list(from = "2010-01-01", to = "2022-01-01"),
  type = c("Federal Election", "By-Election"), # Any or All
  category = "Statistics",
  process = TRUE # OR FALSE
)
```

#### *Sample Data:*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Postal vote applications by party",
               date_range = list(from = "2019-01-01", to = "2022-01-01"),
               category = "Statistics", process = TRUE)
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

### Pre-poll votes
The `Pre-poll votes` dataset provides a summary of the number of prepoll votes issued, detailing:

- the polling place details (`StateAb`, `DivisionNm`, `PollingPlaceNm`).
- the date the pre-poll vote was issued (`IssueDate`).
- the total number of prepoll votes issued (`TotalPPVs`).

Covering elections from `2010` to `2019`, this dataset is available for the `Statistics` category. The dataset is processed in the following way, using the [`process_ppv()`](https://docs.sarahcgall.co.uk/scgElectionsAU/reference/process_ppv.html) function:

- column names are standardised.
- the date columns are pivoted from wide to long and converted into a date format for ease of comparison across years.

#### *Example Usage:*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Pre-poll votes",
  date_range = list(from = "2010-01-01", to = "2022-01-01"),
  type = c("Federal Election", "Referendum", "By-Election"), # Any or All
  category = "Statistics",
  process = TRUE # OR FALSE
)
```

#### *Sample Data:*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Pre-poll votes",
               date_range = list(from = "2019-01-01", to = "2022-01-01"),
               category = "Statistics", process = TRUE)
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

### Votes by SA1
The `Votes by SA1` dataset provides a summary of the number of votes by SA1, detailing:

- the SA1 details (`StateAb`, `DivisionNm`, `PollingPlaceID`, `PollingPlaceNm`, `StatisticalAreaID`).
- the total number of votes (`Votes`).

Covering elections from `2013` to present, this dataset is available for the `Statistics` category. The dataset is processed in the following way, using the [`process_ccd()`](https://docs.sarahcgall.co.uk/scgElectionsAU/reference/process_ccd.html) function:

- column names are standardised and the `year` column is removed.

#### *Example Usage:*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Votes by SA1",
  date_range = list(from = "2013-01-01", to = "2025-01-01"),
  type = c("Federal Election", "Referendum"), # Any or All
  category = "Statistics",
  process = TRUE # OR FALSE
)
```

#### *Sample Data:*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Votes by SA1",
               date_range = list(from = "2019-01-01", to = "2022-01-01"),
               category = "Statistics", process = TRUE)
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>

### Overseas
The `Overseas` dataset provides a summary of the breakdown of overseas voting, detailing:

- the the overseas post (`OverseasPost`).
- the division that the votes were cast (`StateAb`, `DivisionNm`).
- the total number of votes (`TotalVotes`), broken down by postal and prepoll voting types (`PostalVotes`,`PrePollVotes`).

Covering elections from `2013` to present, this dataset is available for the `Statistics` category. The dataset is processed in the following way, using the [`process_overseas()`](https://docs.sarahcgall.co.uk/scgElectionsAU/reference/process_overseas.html) function:

- column names are standardised and the `pp_sort_nm` and `Total` columns are removed.
- the `TotalVotes` column is calculated based on the sum of `PostalVotes` and `PrePollVotes`.

#### *Example Usage:*

```{r, eval = FALSE}
df <- get_election_data(
  file_name = "Overseas",
  date_range = list(from = "2013-01-01", to = "2025-01-01"),
  type = c("Federal Election", "Referendum"), # Any or All
  category = "Statistics",
  process = TRUE # OR FALSE
)
```

#### *Sample Data:*

```{r, echo = FALSE}
df <- suppressMessages(
  get_election_data(file_name = "Overseas",
               date_range = list(from = "2019-01-01", to = "2022-01-01"),
               category = "Statistics", process = TRUE)
)

# Selecting only the first row and keeping all columns
first_row <- df[1,]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>
