---
title: "AEC Federal Election Data"
---
```{=html}
<style>
  .content p {
    margin-top: 30px !important;
  }

  details {
    margin-bottom: 40px;
  }
</style>
```
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
```
<br>
This article explores the [`get_aec_data`](https://sarahcgallLtd.github.io/scgElectionsAU/reference/get_aec_data.html) function and the data that it is able to retrieve from the Australian Electoral Commission's website. This function is primarily retrieving the raw data with limited processing. For a more processed and 'analysis-ready' function, check out `get_XXX`](https://sarahcgallLtd.github.io/scgElectionsAU/reference/get_XXX.html).
<br>

```{r setup, include = FALSE}
library(scgElectionsAU)
library(knitr)
```
In the [`get_aec_data`](https://sarahcgallLtd.github.io/scgElectionsAU/reference/get_aec_data.html) function, the `type` argument default is "Federal Election". This type extracts data from the AEC's 'Tally Room' for elections between 2004 and 2022. Both House and Senate data are extractable.

### National list of candidates
The `National list of candidates` dataset contains a list of candidates including whether or not the candidate was `Elected` or an incumbent (`HistoricElected`). This dataset exists for both `Senate` and `House` candidates.

There is the option to switch `process` to `TRUE`. In doing this, the 2004 dataset will be the same as the other election datasets by removing the `SittingMemberFl` column and replacing it with the `Elected` and `HistoricElected` columns. The `HistoricElected` column is based on the `SittingMemberFl` column and the `Elected` column is based on the `Senators elected` dataset.

<br>
*Options for this dataset include:*

- `file_name = National list of candidates`
- `date_range = list(from = "2004-01-01", to = "2023-01-01")`
- `category = c("House","Senate")`
- `process = TRUE`

<br>
*Dataset output:*

```{r, echo = FALSE}
national_list_of_candidates <- get_aec_data(file_name = "National list of candidates")

# Selecting only the first row and keeping all columns
first_row <- national_list_of_candidates[1, ]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

### Members/Senators elected
The `Members elected` and `Senators elected` datasets list the elected MPs and Senators at each election. The `Senators elected` dataset contains the `ElectedOrder` column.

There is no need to process this data as all election datasets are already standardised.

<br>
*Options for the `Members elected` dataset include:*

- `file_name = Members elected`
- `date_range = list(from = "2004-01-01", to = "2023-01-01")`
- `category = "House"`

<br>
*Dataset output:*

```{r, echo = FALSE}
members_elected <- get_aec_data(file_name = "Members elected")

# Selecting only the first row and keeping all columns
first_row <- members_elected[1, ]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```

<br>
*Options for the `Senators elected` dataset include:*

- `file_name = Senators elected`
- `date_range = list(from = "2004-01-01", to = "2023-01-01")`
- `category = "Senate"`

<br>
*Dataset output:*

```{r, echo = FALSE}
members_elected <- get_aec_data(file_name = "Senators elected", category = "Senate")

# Selecting only the first row and keeping all columns
first_row <- members_elected[1, ]

# Create the HTML table with kable
kable(first_row, format = "html", table.attr = "class='table table-bordered'")
```
